Adaptive Safe Strong Screening for Efficient Lasso-Type Problems Optimization over the Solution Path

### 2020-10-15 main.tex
Updates from Overleaf

\begin{theorem}[Adaptive EDPP]
    For the lasso problem (1), let $\lambda_0:=\max_j|x_j^Ty/n|$. Suppose we have a sequence of $\lambda$ values $\lambda_0>\lambda_1>...>\lambda_K$. Under condition (2)
    \begin{enumerate}
        \item For any $k=1,2,...,K-1$, given $\hat{\beta}(\lambda_k)$, $r(\lambda_k)$ and $\hat{y}(\lambda_k):=X\hat{\beta}(\lambda_k)$, then for some update point $k+B$ and any $b$ s.t. $k<k+b\leq k+B\leq K$, $\hat{\beta}_j(\lambda_{k+b})=0$ if
        \begin{equation}
            \begin{split}
                &\left|2\lambda_{k+b}x_j^Tr(\lambda_k)+(\lambda_k-\lambda_{k+b})\left( x_j^Ty-\frac{y^T\hat{y}(\lambda_k)x_j^T\hat{y}(\lambda_k)}{||\hat{y}(\lambda_k)||_2^2}\right)\right|\\&<2n\lambda_k\lambda_{k+b}-(\lambda_k-\lambda_{k+b})\sqrt{n||y||_2^2-\frac{n(y^T\hat{y}(\lambda_k))^2}{||\hat{y}(\lambda_k)||_2^2}}
            \end{split}
        \end{equation}
        \item For k=0, i.e., $\lambda_k=\lambda_0$, then for some update point $k+B$ and any $b$ s.t. $0<b\leq B\leq K$, $\hat{\beta}_j(\lambda_{k+b})=0$ if
        \begin{equation}
        |(\lambda_0+\lambda_b)x_j^Ty-(\lambda_0-\lambda_b)sign(x_*^Ty)\lambda_0x_j^Tx_*|<2n\lambda_0\lambda_b-(\lambda_0-\lambda_b)\sqrt{n||y||_2^2-n^2\lambda_0^2}.
    \end{equation}
    \end{enumerate}
\end{theorem}

@@

We also tried varying other factors, for example, varying the correlation among features in $X$ or varying the sparsity with fixed signal strength. The AHR also shows uniform advantage over other method when those factors vary in reasonable ranges.

@@

In previous sections, we used simulation to compare the screening methods under different scenario. In real world, data may be generated from different complicated models that cannot be covered by a simulation study, so in this section we are going to test these screening methods on some real data sets to make sure our comparison conclusion is representative in real world. Following the studies of \citep{wang2013lasso, xiang2016screening, zeng2017efficient}, we will test the lasso screening methods on the following data sets:

@@

For 1 and 3, we replicate the experiment 10 times to take the average. For 2 and 4, since in each experiment the response is randomly selected, we replicate the experiment 50 times to get an accurate estimate. The following table compares the average computation time for different screening methods in these data sets.

@@

>>>>

\begin{theorem}[Adaptive EDPP]
    For the lasso problem (1), let $\lambda_0:=\max_j|x_j^Ty/n|$. Suppose we have a sequence of $\lambda$ values $\lambda_0>\lambda_1>...>\lambda_K$ and the last updated reference point is $\lambda_k$. Under condition (2)
    \begin{enumerate}
        \item For any $k=1,2,...,K-1$, given $\hat{\beta}(\lambda_k)$, $r(\lambda_k)$ and $\hat{y}(\lambda_k):=X\hat{\beta}(\lambda_k)$, then for some next update point $k+B$ and any $b$ s.t. $k<k+b\leq k+B\leq K$, $\hat{\beta}_j(\lambda_{k+b})=0$ if
        \begin{equation}
            \begin{split}
                &\left|2\lambda_{k+b}x_j^Tr(\lambda_k)+(\lambda_k-\lambda_{k+b})\left( x_j^Ty-\frac{y^T\hat{y}(\lambda_k)x_j^T\hat{y}(\lambda_k)}{||\hat{y}(\lambda_k)||_2^2}\right)\right|\\&<2n\lambda_k\lambda_{k+b}-(\lambda_k-\lambda_{k+b})\sqrt{n||y||_2^2-\frac{n(y^T\hat{y}(\lambda_k))^2}{||\hat{y}(\lambda_k)||_2^2}}
            \end{split}
        \end{equation}
        \item For k=0, i.e., $\lambda_k=\lambda_0$, then for some next update point $k+B$ and any $b$ s.t. $0<b\leq B\leq K$, $\hat{\beta}_j(\lambda_{k+b})=0$ if
        \begin{equation}
        |(\lambda_0+\lambda_b)x_j^Ty-(\lambda_0-\lambda_b)sign(x_*^Ty)\lambda_0x_j^Tx_*|<2n\lambda_0\lambda_b-(\lambda_0-\lambda_b)\sqrt{n||y||_2^2-n^2\lambda_0^2}.
    \end{equation}
    \end{enumerate}
\end{theorem}

@@

We also tried varying other factors, for example, varying the correlation among features in $X$, varying the sparsity with fixed signal strength or varying the minimum $\lambda$-value $\lambda_K$. The AHR also shows uniform advantage over other method when those factors vary in reasonable ranges.

@@

In previous sections, we used simulation to compare the screening methods under different scenario. In real world, data may be generated from different complicated models that cannot be covered by a simulation study, so in this section we are going to test the screening methods for lasso model on some real data sets to make sure our comparison conclusion is representative in real world. Following the studies of \citep{wang2013lasso, xiang2016screening, zeng2017efficient}, we will test the lasso screening methods on the following data sets:

@@

For 1 and 3, we replicate the experiment 10 times to take the average. For 2 and 4, since in each experiment the response is randomly selected, we replicate the experiment 50 times to get an accurate estimate. Table 1 compares the average computation time for different screening methods in these data sets.

@@

Ultra high dimensional big data that is too large to be fit into the memory is the real challenge and is the place where the help of a good screening method is wanted most. We test the screening methods on a data from a genome-wide association study on 1,080 subjects with heart conditions. We want to identify genetic factors that are associated with some heart condition severity measurements. After imputing and filtering, there are $p=11,830,470$ SNP measurements used as the features. The size of the file storing the feature matrix is 96G and we conduct the test on a machine with 32G RAM.

For the lasso model, we choose the Left ventricular ejection fraction (LVEF) as the response. It measures the proportion of blood being pumped out of the left ventricle of the heart (the main pumping chamber) with each contraction. Lower value means severer heart failure. $n=973$ out of 1,080 subjects have this measurement and they will be considered as the sample for this analysis. The experiment is replicated 5 times to take the average, shown in the Table 2.

\begin{table}[H]
\centering
\begin{tabular}{l|l|l|l|l|l}
\hline

Screening method & AC\,\,\,\,\,\,\,\,\,    & SSR\,\,\,\,\,\,   & SEDPP & HSSR\,\,\,  & AHR\,\,\,\,\,\,  \\
\hline
Time & 256.4 & 79.2 & 79.3 & 84.1 & 54.0\\
Standard error & 0.59 & 0.10 & 0.03 & 0.03 & 0.22
 \\\hline
\end{tabular}
\caption{Average computing time in minute and standard error for solving the lasso model on big data}
\end{table}
